---
title: "Assignment_3"
author: "Alexis Landy"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

```{r global, include=FALSE}
library(readxl)
library(dplyr)
library(tidymodels)
library(tidyverse)
library(packcircles)
library(plotly)
library(viridisLite)
library(fontawesome)
library(shiny)
library(shinyjs)
library(DT)
library(bslib)
library(shinyWidgets)
library(vip)
library(yardstick) 

# ============================================================
# DATA LOADING + CLEANING
# ============================================================

games <- read_excel("C:/Projects/Data-Games/board_games_clean.xlsx")

games_new <- games
games_new[games_new == "NA"] <- NA

games_new$Themes <- toupper(games_new$Theme)
games_new$Mechanic <- toupper(games_new$Category)
games_new$Publishers <- as.factor(games_new$Publisher)

games_analysis <- games_new %>%
  select(1:5, 8, 9, 11:15) %>%
  relocate(Publishers, .after = Years) %>%
  relocate(Themes, .after = Playing_time) %>%
  relocate(Mechanic, .after = Themes) %>%
  rename(
    Play_time = Playing_time,
    Year_published = Years,
    Total_ratings = Number_of_ratings
  )

games_mini <- games_analysis %>% 
  slice(1:1000)

# ============================================================
# DATA SPLITTING
# ============================================================

set.seed(123)
games_splits <- games_analysis %>%
  slice(1:1000) %>%
  select(-NAME) %>%
  initial_split(prop = 0.6, strata = Average_rating)

# ============================================================
# MECHANIC DICTIONARY
# ============================================================

mechanic_dictionary <- tibble::tribble(
  ~Mechanic, ~Definition, ~Examples,
  "Dice Rolling",
  "A mechanic where players roll dice to generate random outcomes that influence actions, resources, or combat.",
  "Yahtzee (crowd), King of Tokyo (expert)",
  
  "Roll / Spin and Move",
  "A mechanic where players roll dice or spin a wheel to determine how far they move along a track or board.",
  "Monopoly (crowd), Talisman (expert)",
  
  "Chit‑Pull System",
  "A mechanic where random tokens are drawn from a pool to determine events, actions, or combat resolution.",
  "War Chest (crowd), Combat Commander (expert)",
  
  "Press Your Luck",
  "A mechanic where players repeatedly take risky actions for increasing rewards, stopping before they bust.",
  "Can't Stop (crowd), Quacks of Quedlinburg (expert)",
  
  "Rock‑Paper‑Scissors",
  "A simultaneous choice mechanic where players select actions that resolve through a cyclical dominance relationship.",
  "Kingdomino Duel (crowd), Inis (expert)",
  
  "Hand Management",
  "A mechanic where players optimize the timing and combination of cards or tokens in their hand to maximize value.",
  "Uno (crowd), Race for the Galaxy (expert)",
  
  "Set Collection",
  "A mechanic where players gather groups of related items to score points or unlock abilities.",
  "Rummy (crowd), Sushi Go! (expert)",
  
  "Card Drafting",
  "A mechanic where players choose cards from a shared pool, often passing the remaining cards to others.",
  "7 Wonders (crowd), Blood Rage (expert)",
  
  "Deck / Pool Building",
  "A mechanic where players improve a personal deck or pool of actions by acquiring new, stronger cards.",
  "Dominion (crowd), Aeon’s End (expert)",
  
  "Action Queue",
  "A mechanic where players pre‑arrange actions in a sequence that resolves later, often with limited ability to change it.",
  "RoboRally (crowd), Mechs vs. Minions (expert)",
  
  "Action Points",
  "A mechanic where players spend a limited pool of points each turn to perform actions of varying costs.",
  "Pandemic (crowd), Tikal (expert)",
  
  "Variable Phase Order",
  "A mechanic where the sequence of game phases changes from round to round based on player choices or effects.",
  "Small World (crowd), Puerto Rico (expert)",
  
  "Variable Player Powers",
  "A mechanic where each player has unique abilities or advantages that shape their strategy and playstyle.",
  "King of Tokyo (crowd), Root (expert)",
  
  "Commodity Speculation",
  "A mechanic where players buy, sell, or hold goods whose values fluctuate over time.",
  "Pit (crowd), Container (expert)",
  
  "Stock Holding",
  "A mechanic where players acquire shares in companies or factions to earn dividends or influence outcomes.",
  "Acquire (crowd), Imperial 2030 (expert)",
  
  "Grid Movement",
  "A mechanic where players move pieces across a board divided into discrete squares or spaces.",
  "Chess (crowd), Gloomhaven (expert)",
  
  "Area Movement",
  "A mechanic where players move between large, connected regions rather than fixed grid spaces.",
  "Risk (crowd), Twilight Imperium (expert)",
  
  "Point‑to‑Point Movement",
  "A mechanic where movement occurs along predefined paths connecting specific nodes on a map.",
  "Ticket to Ride (crowd), War of the Ring (expert)",
  
  "Hex‑and‑Counter",
  "A wargame mechanic where units move and fight on a hex‑based map for precise tactical positioning.",
  "BattleLore (crowd), Advanced Squad Leader (expert)",
  
  "Hexagon Grid",
  "A mechanic where movement and adjacency are defined by hexagonal spaces.",
  "Catan (crowd), Mage Knight (expert)",
  
  "Modular Board",
  "A mechanic where the board is built from separate tiles or sections that change layout each game.",
  "Carcassonne (crowd), Eclipse (expert)",
  
  "Tile Placement",
  "A mechanic where players place tiles to build patterns, score points, or shape the board.",
  "Carcassonne (crowd), Azul (expert)",
  
  "Pattern Building",
  "A mechanic where players create specific arrangements of pieces or symbols to score points.",
  "Qwirkle (crowd), Sagrada (expert)",
  
  "Pattern Recognition",
  "A mechanic where players identify or react to visual patterns quickly or accurately.",
  "Spot It! (crowd), The Mind (expert)",
  
  "Line Drawing",
  "A mechanic where players draw lines or routes to claim areas, connect points, or fulfill objectives.",
  "Dots and Boxes (crowd), Cartographers (expert)",
  
  "Area Control / Majority",
  "A mechanic where players compete to dominate regions by having the most presence or influence.",
  "Risk (crowd), El Grande (expert)",
  
  "Area Enclosure",
  "A mechanic where players create closed boundaries on the board to claim territory.",
  "Go (crowd), Qin (expert)",
  
  "Area‑Impulse",
  "A mechanic where players activate areas one at a time to move or attack within those zones.",
  "Nexus Ops (crowd), Storm Over Arnhem (expert)",
  
  "Route / Network Building",
  "A mechanic where players construct connected paths or networks to link locations or deliver goods.",
  "Ticket to Ride (crowd), Power Grid (expert)",
  
  "Action / Movement Programming",
  "A mechanic where players secretly plan actions in advance, revealing them later in a fixed sequence.",
  "RoboRally (crowd), Colt Express (expert)",
  
  "Simultaneous Action Selection",
  "A mechanic where all players choose actions at the same time and reveal them together.",
  "Wits & Wagers (crowd), 7 Wonders Duel (expert)",
  
  "Trick‑Taking",
  "A mechanic where players play cards into a trick, with one card winning based on suit or rank rules.",
  "Hearts (crowd), The Crew (expert)",
  
  "Memory",
  "A mechanic where players rely on remembering hidden information or past actions.",
  "Memory Match (crowd), Mysterium (expert)",
  
  "Storytelling",
  "A mechanic where players create or interpret narratives to guide play or communicate clues.",
  "Dixit (crowd), Once Upon a Time (expert)",
  
  "Acting",
  "A mechanic where players perform gestures or movements to convey information or complete tasks.",
  "Charades (crowd), Time’s Up! (expert)",
  
  "Cooperative Play",
  "A mechanic where players work together toward a shared objective, winning or losing as a team.",
  "Pandemic (crowd), Spirit Island (expert)",
  
  "Partnerships",
  "A mechanic where players form fixed or temporary teams to coordinate strategies.",
  "Spades (crowd), The Crew: Mission Deep Sea (expert)",
  
  "Team‑Based Game",
  "A mechanic where players are divided into opposing teams that coordinate internally while competing externally.",
  "Codenames (crowd), Captain Sonar (expert)",
  
  "Negotiation",
  "A mechanic where players bargain or persuade others to gain advantages or influence outcomes.",
  "Monopoly (crowd), Chinatown (expert)",
  
  "Alliances",
  "A mechanic where players form agreements to cooperate temporarily for mutual benefit.",
  "Risk (crowd), Diplomacy (expert)",
  
  "Bribery",
  "A mechanic where players offer incentives or resources to influence others’ decisions.",
  "Sheriff of Nottingham (crowd), Republic of Rome (expert)",
  
  "Voting",
  "A mechanic where players cast votes to decide outcomes or trigger events.",
  "The Resistance (crowd), Dead of Winter (expert)",
  
  "Take That",
  "A mechanic where players directly hinder opponents through attacks or disruptive actions.",
  "Exploding Kittens (crowd), Munchkin (expert)",
  
  "Secret Unit Deployment",
  "A mechanic where players place units or choose actions secretly, revealing them later.",
  "Stratego (crowd), Cry Havoc (expert)",
  
  "Player Elimination",
  "A mechanic where players can be removed from the game before it ends.",
  "Clue (crowd), Bang! (expert)",
  
  "Pick‑Up and Deliver",
  "A mechanic where players transport goods or characters between locations to fulfill objectives.",
  "Ticket to Ride: First Journey (crowd), Merchants & Marauders (expert)",
  
  "Trading",
  "A mechanic where players exchange resources or favors to improve their position.",
  "Catan (crowd), Bohnanza (expert)",
  
  "Simulation",
  "A mechanic where the game models real‑world systems or scenarios with interconnected variables.",
  "SimCity Board Game (crowd), Food Chain Magnate (expert)",
  
  "Time Track",
  "A mechanic where turn order is determined by how much time players spend on actions.",
  "Patchwork (crowd), The Red Cathedral (expert)",
  
  "Real‑Time",
  "A mechanic where players act simultaneously under time pressure instead of taking turns.",
  "5‑Minute Dungeon (crowd), Space Alert (expert)",
  
  "Role Playing",
  "A mechanic where players assume characters with defined traits and make decisions based on narrative identity.",
  "Dungeons & Dragons (crowd), Gloomhaven: Jaws of the Lion (expert)",
  
  "Campaign / Battle Card Driven",
  "A mechanic where scenarios or battles progress through card‑driven events or story developments.",
  "Risk Legacy (crowd), Twilight Struggle (expert)",
  
  "Wargames",
  "A mechanic category where players simulate military conflict using units, terrain, and tactical decisions.",
  "Axis & Allies (crowd), Advanced Squad Leader (expert)",
  
  "Solo / Solitaire Game",
  "A mechanic where the game is designed for a single player using automated opponents or puzzles.",
  "Friday (crowd), Mage Knight Solo (expert)",
  
  "Children’s Games",
  "A mechanic category where rules and components are simplified for younger players.",
  "Candy Land (crowd), Outfoxed! (expert)",
  
  "Singing",
  "A mechanic where players must sing or vocalize as part of performing actions or giving clues.",
  "Happy Salmon (crowd), Spontuneous (expert)")

# ============================================================
# TOP 10 MECHANICS PREP
# ============================================================

mechanic_counts <- games_analysis %>%
  separate_rows(Mechanic, sep = ";") %>%
  mutate(Mechanic = str_trim(Mechanic)) %>%
  count(Mechanic, sort = TRUE)

# ============================================================
# MODELING PIPELINE
# ============================================================

games_recipe <-
  recipe(Average_rating ~ ., data = training(games_splits)) %>%
  step_naomit() %>% 
  step_unknown(all_nominal_predictors()) %>% 
  step_impute_mean(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())

games_glmnet <- 
  linear_reg(penalty = tune(), mixture = 1) %>%   # mixture = 1 = lasso
  set_engine("glmnet")

glmnet_wflow <- 
  workflow() %>%
  add_model(games_glmnet) %>%
  add_recipe(games_recipe)

set.seed(123)
games_folds <- vfold_cv(training(games_splits), v = 5)

glmnet_grid <- tibble(
  penalty = 10^seq(-4, -1, length.out = 20)
)

# ============================================================
# CIRCLE PACKING
# ============================================================

clusters <- split(games_analysis, games_analysis$Mechanic)

sampled_clusters <- lapply(clusters, function(df) {
  if (nrow(df) <= 15) df else df %>% slice_sample(n = 15)
})

truncate_chars <- function(x, limit = 15) {
  sapply(x, function(name) {
    if (nchar(name) <= limit) return(name)
    truncated <- substr(name, 1, limit)
    truncated <- sub("\\s+\\S*$", "", truncated)
    paste0(truncated, "...")
  })
}

games_sampled <- bind_rows(sampled_clusters)

radius <- log1p(games_sampled$Play_time)

games_sampled_final <- games_sampled %>% 
  mutate(Mechanic = as.character(Mechanic),
         Radius   = radius)

games_sampled_split <- split(games_sampled_final, games_sampled_final$Mechanic)

split_clusters <- lapply(games_sampled_split, function(df){
  mech <- df$Radius
  packing_split <- circleLayout(mech)$layout %>%
    rename(pack_x = x, pack_y = y, pack_r = radius)
  cbind(df, packing_split)
})

games_packed_all <- bind_rows(split_clusters)

scale_factor <- 300

games_packed_all <- games_packed_all %>%
  group_by(Mechanic) %>%
  mutate(pack_x = pack_x * scale_factor,
         pack_y = pack_y * scale_factor) %>%
  mutate(pack_x = pack_x - mean(pack_x),
         pack_y = pack_y - mean(pack_y)) %>%
  ungroup()

clusters_per_row <- ceiling(60 / 13)

Mech_index <- games_packed_all %>%
  distinct(Mechanic) %>%
  arrange(Mechanic) %>%
  mutate(Index = row_number(),
         Row   = (Index - 1) %/% clusters_per_row,
         Col   = (Index - 1) %% clusters_per_row)

games_packed_final <- games_packed_all %>%
  left_join(Mech_index, by = "Mechanic") %>%
  mutate(Name_short  = truncate_chars(NAME, 15),
         x_shifted   = pack_x + Col * 35000,
         y_shifted   = pack_y - Row * 20000,
         Themes_clean = as.character(Themes))

mech_list <- sort(unique(games_packed_final$Mechanic))

facet1_mechs <- mech_list[1:33]
facet2_mechs <- mech_list[34:65]

facet1_data <- games_packed_final %>% filter(Mechanic %in% facet1_mechs)
facet2_data <- games_packed_final %>% filter(Mechanic %in% facet2_mechs)

theme_colors <- viridisLite::turbo(length(unique(games_packed_final$Themes_clean)))

circles <- function(df) {
  plot_ly(
    data = df,
    type = "scatter",
    mode = "markers",
    x = ~x_shifted,
    y = ~y_shifted,
    color = ~Themes_clean,
    colors = theme_colors,
    marker = list(
      sizemode = "diameter",
      size     = ~pack_r * 5,
      alpha    = 0.7
    ),
    text = ~paste0(
      "<b>", Name_short, "</b><br>",
      "Mechanic: ", Mechanic, "<br>",
      "Playtime (min): ", Play_time, "<br>",
      "Publisher: ", Publishers
    ),
    hoverinfo = "text"
  )
}

```


---

Sidebar {.sidebar}
===========================================================

<br>

`r h3("Objective:")`
`r h3("Predict which hotel stays include children.")`

<br>

- - -

<br

```{r}
useShinyjs()

selectInput("select_mechanic", "Mechanics:", choices = sort(unique(games_analysis$Mechanic)), multiple = TRUE)

sliderInput("slide_player", "Minimum Player Number", min = 2, max = 9, value = 2)

sliderInput("slide_time", "Playtime (Min)", min = 5, max = 60000, value = 30)

selectInput("select_theme", "Themes:", choices = sort(unique(games_analysis$Themes)), multiple = TRUE)

actionButton("apply", "Apply", icon = icon("circle-play"))
actionButton("reset", "Reset", icon = icon("arrow-rotate-left"))

```

```{r}
observeEvent(input$reset, {
  updateSelectInput(session, "select_mechanic", selected = character(0))
  updateSelectInput(session, "select_theme",   selected = character(0))
  updateSliderInput(session, "slide_player",   value = 2)
  updateSliderInput(session, "slide_time",     value = 30)

  shinyjs::delay(300, {
    shinyjs::click("apply")
  })
})

games_filtered <- eventReactive(input$apply, {
  df <- games_analysis

  if (length(input$select_mechanic) > 0)
    df <- df %>% filter(Mechanic %in% input$select_mechanic)

  if (length(input$select_theme) > 0)
    df <- df %>% filter(Themes %in% input$select_theme)

  df %>%
    filter(
      Min_players <= input$slide_player,
      Play_time   <= input$slide_time
    )
}, ignoreNULL = FALSE)

```


Game Comparison
===========================================================

### Game Comparison

```{r, include=FALSE}
glmnet_tuned <- tune_grid(
  glmnet_wflow,
  resamples = games_folds,
  grid = glmnet_grid,
  metrics = metric_set(rmse)
)

best_glmnet <- select_best(glmnet_tuned, metric = "rmse")

final_glmnet_fit <- 
  finalize_workflow(glmnet_wflow, best_glmnet) %>%
  fit(training(games_splits))

final_glmnet_fit %>% 
  predict(training(games_splits))

final_glmnet_fit %>%
  predict(testing(games_splits)) %>%
  bind_cols(testing(games_splits)) %>%
  metrics(truth = Average_rating, estimate = .pred)

```


## Mechanics A–M
```{r}
output$plot_1 <- renderPlotly({
  circles(facet1_data)
})

plotlyOutput("plot_1")
```

## Mechanics N–Z
```{r}
output$plot_2 <- renderPlotly({
  circles(facet2_data)
})

plotlyOutput("plot_2")

```


---

Find a Game
===========================================================
### Games Table

## boon
```{r}
output$games_table_final <- renderDT({
  datatable(games_filtered())
})

DTOutput("games_table_final")

```

## Rating Box
```{r}
output$rating_box <- renderText({
  avg_rating <- mean(games_filtered()$Average_rating, na.rm = TRUE)
  ifelse(is.na(avg_rating), "NA", paste0(round(avg_rating, 1), "/10"))
})

```

```{css}
.kpi-card {
  background-color: #6a0dad;
  color: white;
  padding: 25px;
  border-radius: 12px;
  font-size: 32px;
  font-weight: bold;
  text-align: center;
  margin-top: 10px;
  margin-bottom: 20px;
}
```

<div class="kpi-card">
  ```{r}
  textOutput("rating_box")
  ```
</div>

### Top Mechanics
```{r}
games_analysis %>%
  separate_rows(Mechanic, sep = ";") %>%
  mutate(Mechanic = str_trim(Mechanic)) %>%
  count(Mechanic, sort = TRUE) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(Mechanic, n), y = n)) +
  geom_col() +
  coord_flip()
```

---

Model Performance
===========================================================
### Data Splitting

**Total Observations:**  
`r dim(games_analysis)%>% scales::comma()`

**Training Set:**  
`r dim(training(games_splits)) %>% scales::comma()`

**Validation Set:**  
`r (dim(training(games_splits))* 0.6) %>% scales::comma()`

**Testing Set:**  
`r dim(testing(games_splits))%>% scales::comma()`


## Model Accuracy (linear regression)
```{r}
final_glmnet_fit %>% 
  augment(new_data = testing(games_splits)) %>% 
  ggplot(aes(x = .pred, y = Average_rating)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)

```

## Prediction Errors
```{r}
final_glmnet_fit %>%
  predict(new_data = testing(games_splits)) %>%
  bind_cols(testing(games_splits)) %>%
  metrics(truth = Average_rating, estimate = .pred) %>% 
  ggplot(aes(x = .metric, y = .estimate, fill = .metric)) +
  geom_col(width = 0.6, alpha = 0.8) +
  scale_fill_manual(values = c("rmse" = "#4B9CD3",
                               "mae"  = "#7EC8E3",
                               "rsq"  = "#1B4965")) +
  labs(
    title = "Model Prediction Error Metrics",
    x = "Metric",
    y = "Value"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

```


## Prediction Metrics
```{r}
final_glmnet_fit %>%
  predict(testing(games_splits)) %>%
  bind_cols(testing(games_splits)) %>%
  metrics(truth = Average_rating, estimate = .pred)

```

## VIP Plot
```{r}
final_glmnet_fit %>% 
  extract_fit_parsnip() %>%
  vip::vip() +
  geom_col(aes(fill = Variable))
```


---

Data Overview
===========================================================
### Games Table

```{r}
output$games_table_final <- renderDT({
  datatable(
    games_analysis %>% slice(1:1000),
    filter = "top",
    options = list(
      pageLength = 100,
      scrollY = "400px",
      scrollX = TRUE
    )
  )
})
```

```{r}
DTOutput("games_table_final")
```

## Dictionary
```{r}
output$mechanic_table <- renderDT({
  mechanic_dictionary %>%
    arrange(Mechanic) %>%
    datatable(
      options = list(
        searching = TRUE,
        pageLength = 50,
        lengthMenu = c(50, 100),
        scrollY = "250px",
        scrollX = "300px"
      ),
      style = "default"
    )
})

DTOutput("mechanic_table")
```
